Process RNA-Seq and Genotype Data for TWAS

### 2019-10-16 code/ProcessData.Rmd
Annotate MSBB By Brain Region

```{bash PLINKSetup, results='hide', eval=F, echo=TRUE}
mkdir ~/TWAS/bin
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink1-assets/plink_linux_x86_64_20190617.zip 
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink_linux_x86_64_20190617.zip
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20191006.zip
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink2_linux_x86_64_20191006.zip

@@

## MSBB RNA-Seq: Average Across Brain Regions, Filter for INDV. w/Genotype Data, Rename Samples with Genotype ID 
```{r MSBBRNA, results='hide', echo=TRUE, cache = FALSE}
library(data.table)

Process_MSBB <- function( dna, exp, id){
  #'@dna synID of Geno samples eg 'syn11384608'
  #'@exp synID of expression data eg 'syn8485026'
  #'@id synID of id translators  eg. 'syn6100548'

  DNA <- fread(syn_temp$get(dna)$path)
  Expression <- read.table(syn_temp$get(exp)$path, header=T)
  IDs <- fread(syn_temp$get(id)$path)
  IDs <- IDs[ grepl( 'accepted_hits.sort.coord.bam', IDs$fileName ) == T, ]
  IDs <- IDs[ (IDs$sampleIdentifier %in% colnames(Expression)) == T, ]

  INDVs <- IDs$individualIdentifier[ !duplicated(IDs$individualIdentifier) ]

  if( 'ensembl_gene_id' %in% colnames(Expression) ){
    Averaged <- as.data.frame( Expression$ensembl_gene_id )
  }else{
    Averaged <- as.data.frame( row.names(Expression) )
    colnames(Averaged) <- 'ensembl_gene_id'
  }

  Ave <- function( ID, DF, Exp ){
    #'@ID Sample ID eg. 'AMPAD_MSSM_0000007155'
    #'@DF the target dataframe eg. 'IDs'
    #'@Exp the expression set to filter eg Expression

    Exp <- as.data.frame( Exp )
    COLS <- DF[ DF$individualIdentifier == ID, ]$sampleIdentifier
    Small <- Exp[ , COLS ]

    if( dim(as.matrix(Small))[2] > 1 ){
      eval(parse( text = paste0( 'return( ', ID, ' = rowMeans(Small) )')))
    }else{
      eval(parse( text = paste0( 'return( ', ID, ' = Small)')))
    }
  }

  for( i in INDVs ){
    eval(parse( text = paste0( 'Averaged$', i, ' <- Ave(i, IDs, Expression )') ))
  }

  GWAS_ID <- as.data.frame( DNA[ (DNA$individualIdentifier %in% colnames(Averaged)) == T , ] )
  row.names(GWAS_ID) <- GWAS_ID$individualIdentifier
  if( 'ensembl_gene_id' %in% colnames(Averaged) ){
    Averaged <- Averaged[ , c( 'ensembl_gene_id', row.names(GWAS_ID) )]
  }else{
    Averaged <- Averaged[ , c( 'Expression$ensembl_gene_id', row.names(GWAS_ID) )]
    colnames(Averaged[1]) <- 'ensembl_gene_id'
  }
  colnames(Averaged) <- paste0(GWAS_ID[ colnames(Averaged), ]$sampleIdentifier, "_", GWAS_ID[ colnames(Averaged), ]$sampleIdentifier)
  colnames(Averaged)[1] <- "ensembl_gene_id"
  return( Averaged )
}

@@

  fam <- FAM
  if( Type == 0){
    RNA <- read.table( syn_temp$get(syn)$path)
    colnames(RNA) <- gsub('_TCX', '_a', colnames(RNA))
    colnames(RNA) <- gsub('X', '', colnames(RNA))
    colnames(RNA) <- gsub('_a', '_TCX', colnames(RNA))
    RNA <- cbind( 'ensembl_gene_id' = row.names(RNA), RNA[,colnames(RNA)] )
  }else{
    RNA <- as.data.frame( fread( syn_temp$get(syn)$path ) )
  }
  RNA <- RNA[ , c('ensembl_gene_id', colnames(RNA)[(grepl('_TCX', colnames(RNA))==T)]) ]
  colnames( RNA ) <- gsub( '_TCX', '', colnames( RNA ) )
  colnames( RNA )[2:length(colnames( RNA ))] <- paste0( '0_', colnames( RNA )[2:length(colnames( RNA ))] )

  RNA <- RNA[ , c( "ensembl_gene_id", colnames(RNA)[ (colnames(RNA) %in% paste0( FAM[,1], "_", FAM[,2] )) == T ]) ]
  return(RNA)
}
Norm_Resid_MAYO <- Mayo_filt( 'syn8466825', FAM, 1 )
dCorrect_Resid_MAYO <- Mayo_filt( 'syn20801991', FAM, 0 )
```

@@

  #Filter for matched RNA-Seq and GWAS samples
  Exp <- Exp[ ,c('ensembl_gene_id', colnames(Exp)[ (colnames(Exp) %in% Trans$rnaseq_id) == T]) ]
  Trans <- as.data.frame( Trans[ (Trans$rnaseq_id %in% colnames(Exp)) == T,  ] )
  row.names(Trans) <- Trans$rnaseq_id

  colnames(Exp) <- c('ensembl_gene_id', paste0( Trans[colnames(Exp)[2:length(colnames(Exp))],]$gwas_id, "_",  Trans[colnames(Exp)[2:length(colnames(Exp))],]$gwas_id))
  return(Exp)
}

@@

## Merge RNA-Seq: Merge and Annotate RNA-Seq
* Remove genes on Chr X, Chr Y, Chr M 
* Check for no Patch Contig Nonsense
* Print Sample ID List
* Filter Genotypes for RNA Samples
```{r Merge RNASeq, results='hide', echo=TRUE, cache = FALSE}
#Common Genes:
GNames <- names( table(c( as.character(Norm_Resid_MSBB$ensembl_gene_id), as.character(Norm_Resid_MAYO$ensembl_gene_id), as.character(Norm_Resid_RM$ensembl_gene_id)  ))[ table(c( as.character(Norm_Resid_MSBB$ensembl_gene_id), as.character(Norm_Resid_MAYO$ensembl_gene_id), as.character(Norm_Resid_RM$ensembl_gene_id)  )) == 3] )

@@

  row.names(MSBB) <- as.character(MSBB$ensembl_gene_id)
  row.names(Mayo) <- as.character(Mayo$ensembl_gene_id)
  row.names(RM) <- as.character(RM$ensembl_gene_id)

@@

#Filter out the ~ 480 ChrX genes
Norm_Resid <- Norm_Resid[ (Norm_Resid$Chr %in% 'X') == F, ]
Diag_Resid <- Diag_Resid[ (Diag_Resid$Chr %in% 'X') == F, ]

@@

#Write New FAM to file
write.table(FAM, file='New_Fam.fam', row.names = F, col.names=F, quote=F, sep=" ")
```

## Filter Genotype data
```{bash FilterGenos, results='hide', echo=TRUE, cache=TRUE}
#source /root/.bashrc
PATH=$PATH:~/TWAS/bin/
CORES=`expr $(nproc) - 2`

@@

>>>>

```{bash PLINKSetup, results='hide', eval=F, echo=TRUE}
mkdir ~/TWAS/bin
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink1-assets/plink_linux_x86_64_20190617.zip 
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink_linux_x86_64_20190617.zip
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20191012.zip
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink2_linux_x86_64_20191012.zip

@@

## MSBB RNA-Seq: Label Brain Regions, Filter for INDV. w/Genotype Data, Rename Samples with Genotype ID 
```{r MSBBRNA, results='hide', echo=TRUE, cache = FALSE}
library(data.table)

Process_MSBB <- function( dna, exp, id){
  #'@dna synID of Geno samples eg 'syn11384608'
  #'@exp synID of expression data eg 'syn8485026'
  #'@id synID of RNA id translators  eg. 'syn6100548'

  DNA <- fread(syn_temp$get(dna)$path)
  Expression <- read.table(syn_temp$get(exp)$path, header=T, row.names = 1)
  IDs <- fread(syn_temp$get(id)$path)
  sink <- Expression

  #Make sample IDs the row names
  IDs <- IDs[ IDs$Action != 'Exclude', ]
    ###Convert to Z-Value by Tissue
  Regions <- names( table(IDs$BrodmannArea) ) 
  for( BM in Regions ){
    temp <- Expression[ ,colnames(Expression)[ colnames(Expression) %in% IDs[ IDs$BrodmannArea == BM, ]$sampleIdentifier] ]
    foo <- apply(temp, 1, scale)
    foo <- t(foo)
    colnames(foo) <- colnames(temp)
    Expression[ , colnames(foo)] <- foo [, colnames(foo)]
  }

  IDs <- IDs[ (IDs$sampleIdentifier %in% colnames(Expression)) == T, ]
  Expression <- Expression[ , colnames(Expression)[(colnames(Expression) %in% IDs$sampleIdentifier) == T] ]
  INDVs <- IDs[ IDs$fileType=='fastq' , ]

  DNA <- as.data.frame(DNA)
  DNA[ is.na(DNA$individualIdentifier)==T,]$individualIdentifier <- DNA[ is.na(DNA$individualIdentifier)==T,]$individualIdentifier.inferred
  DNA <- DNA[ complete.cases(DNA$individualIdentifier), ]
  DNA <- DNA[ DNA$Action != "Exclude", ]
  row.names(DNA) <- DNA$individualIdentifier

  INDVs <- IDs[ !duplicated(IDs$sampleIdentifier) ]
  INDVs <- INDVs[ (INDVs$individualIdentifier %in% DNA$individualIdentifier) == T, ]

  DNA <- as.data.frame(DNA)
  DNA <- DNA[ complete.cases(DNA$individualIdentifier), ]
  row.names(DNA) <- DNA$individualIdentifier

  INDVs <- as.data.frame(INDVs)
  INDVs$GenoID <- DNA[ INDVs$individualIdentifier, ]$sampleIdentifier
  INDVs$NewID <- paste0(INDVs$GenoID, '_', INDVs$BrodmannArea )
  row.names(INDVs) <- INDVs$sampleIdentifier
  INDVs[ colnames(Expression), ]$NewID

  Expression <- Expression[ , colnames(Expression)[(colnames(Expression) %in% INDVs$sampleIdentifier) == T] ]
  INDVs <- as.data.frame(INDVs)
  row.names(INDVs) <- INDVs$sampleIdentifier
  colnames(Expression) <- INDVs[ colnames(Expression), ]$NewID

  return( Expression )
}

@@

  #syn <- 'syn20801991'
  #fam <- FAM
  #Type <- 1

  fam <- FAM
  if( Type == 0){
    RNA <- read.table( syn_temp$get(syn)$path)
    colnames(RNA) <- gsub('_TCX', '_a', colnames(RNA))
    colnames(RNA) <- gsub('X', '', colnames(RNA))
    colnames(RNA) <- gsub('_a', '_TCX', colnames(RNA))
    RNA <- cbind( 'ensembl_gene_id' = row.names(RNA), RNA[,colnames(RNA)] )
    RNA <- RNA[ , colnames(RNA)[(grepl('_TCX', colnames(RNA))==T)] ]
  }else{
    RNA <- as.data.frame( fread( syn_temp$get(syn)$path ) )
    row.names(RNA) <- RNA$ensembl_gene_id
    RNA <- RNA[ , colnames(RNA)[(grepl('_TCX', colnames(RNA))==T)] ]
  }
  colnames( RNA ) <- gsub( '_TCX', '', colnames( RNA ) )
  colnames( RNA ) <- paste0( '0_', colnames( RNA ) )

  ## Z-Scale
  foo <- as.matrix(RNA)
  foo <- apply(RNA, 1, scale)
  foo <- t(foo)
  colnames(foo) <- colnames(RNA)
  #Expression[ , colnames(foo)] <- foo [, colnames(foo)]

  RNA <- foo[ , colnames(foo)[ (colnames(foo) %in% paste0( FAM[,1], "_", FAM[,2] )) == T ] ]
  return(RNA)
}
Norm_Resid_MAYO <- Mayo_filt( 'syn8466825', FAM, 1 )
dCorrect_Resid_MAYO <- Mayo_filt( 'syn20801991', FAM, 0 )
```

@@

  ## Z-Scale
  row.names( Exp ) <- Exp$ensembl_gene_id
  Exp <- Exp[ , (colnames( Exp ) %in% "ensembl_gene_id") == F ]
  foo <- as.matrix(Exp)
  foo <- apply(Exp, 1, scale)
  foo <- t(foo)
  colnames(foo) <- colnames(Exp)

  #Filter for matched RNA-Seq and GWAS samples
  Exp <- foo[ , colnames(foo)[ (colnames(foo) %in% Trans$rnaseq_id) == T] ]
  Trans <- as.data.frame( Trans[ (Trans$rnaseq_id %in% colnames(Exp)) == T,  ] )
  row.names(Trans) <- Trans$rnaseq_id

  colnames(Exp) <- c( paste0( Trans[colnames(Exp)[1:length(colnames(Exp))],]$gwas_id, "_",  Trans[colnames(Exp)[1:length(colnames(Exp))],]$gwas_id))
  return(Exp)
}

@@

## Merge RNA-Seq: Merge and Annotate RNA-Seq
* Remove genes on Chr X, Chr Y, Chr M 
* Check for no Patch Contig Nonsense
* Print Sample ID List
* Filter Genotypes for RNA Samples
```{r Merge RNASeq, results='hide', echo=TRUE, cache = FALSE}
#Common Genes:
GNames <- names( table(c( row.names(Norm_Resid_MSBB), row.names(Norm_Resid_MAYO), row.names(Norm_Resid_RM)  ))[ table(c( row.names(Norm_Resid_MSBB), row.names(Norm_Resid_MAYO), row.names(Norm_Resid_RM)  )) == 3] )

@@

  #row.names(MSBB) <- as.character(MSBB$ensembl_gene_id)
  #row.names(Mayo) <- as.character(Mayo$ensembl_gene_id)
  #row.names(RM) <- as.character(RM$ensembl_gene_id)

@@

#Filter out the ~ ChrX, MT, and Patch genes
Norm_Resid <- Norm_Resid[ (as.character(Norm_Resid$Chr) %in% c('MT', 'X', as.character(Norm_Resid$Chr)[ grepl('PATCH', as.character(Norm_Resid$Chr)) == T])) == F, ]
Diag_Resid <- Diag_Resid[ (as.character(Diag_Resid$Chr) %in% c('MT', 'X', as.character(Diag_Resid$Chr)[ grepl('PATCH', as.character(Diag_Resid$Chr)) == T])) == F, ]

@@

#Write New FAM to file
write.table(FAM, file='New_Fam.fam', row.names = F, col.names=F, quote=F, sep=" ")

tail(FAM[ FAM$V1[ grepl('ROS', FAM$V1) == F ] $ FAM$V1[ grepl('MAP', FAM$V1) == F ] ,])
# & FAM$V1[ grepl('ROS', FAM$V1) == F ]
MSBB_List <- FAM[ FAM$V1 != '0', ] 
MSBB_List <- MSBB_List[ MSBB_List$V1[ grepl('MAP', MSBB_List$V1) == F ] ,]
MSBB_List <- MSBB_List[ MSBB_List$V1[ grepl('ROS', MSBB_List$V1) == F ] ,]
write.table( MSBB_List[,1:2], file="MSBB_SamplesToExpand.txt", quote = F, row.names = F, col.names = F, sep = "\t" )
```

## Filter Genotype data
```{bash FilterGenos, results='hide', echo=TRUE, cache=TRUE}
#source /root/.bashrc
export PATH=$PATH:~/TWAS/bin/
CORES=`expr $(nproc) - 2`

@@

#Extract the MSBB Samples to duplicate and rename to match tissues
plink --threads $CORES --bfile CEU_AncestryMatched_Genotypes --keep MSBB_SamplesToExpand.txt --recode --out MSBB_Intermediate

#Repeat the rows in the ped file 4 times
awk '{for(i=0;i<4;i++)print}' MSBB_Intermediate.ped > temp

gawk -v m=1 -v n=2 -v el='BM10' 'NR % m == 0 { $n = el } 1' temp > foo
gawk -v m=2 -v n=2 -v el='BM22' 'NR % m == 0 { $n = el } 1' foo > temp
gawk -v m=3 -v n=2 -v el='BM36' 'NR % m == 0 { $n = el } 1' temp > foo
gawk -v m=4 -v n=2 -v el='BM44' 'NR % m == 0 { $n = el } 1' foo > temp
mv temp MSBB_Intermediate.ped
rm foo

plink --threads $CORES --file MSBB_Intermediate --make-bed --out MSBB_Expanded
plink --threads $CORES --bfile CEU_AncestryMatched_Genotypes --bmerge MSBB_Expanded --make-bed --out Rebuilt
plink --threads $CORES --bfile Rebuilt --make-bed --out CEU_AncestryMatched_Genotypes

rm Rebuilt.*
rm MSBB_Expanded.*
